Synpotic Soil Characterization

### 2025-09-08 3-reports/manuscript_figures.Rmd
update PCA
relabel wetland --> marsh

```{r targets}
targets::tar_load(c(data_combined, sample_key))
```

@@

 # data_combined_subset = 
    data_combined %>% 
    filter(!transect %in% "wte") %>% 
    force()
}
make_data_wide = function(data_combined_subset, sample_key){

@@

  grp = 
    dat %>% 
    dplyr::select(where(is.character)) %>% 
    dplyr::mutate(row = row_number()) %>% 
    right_join(num_row_numbers)

@@

```{r pca_input}
pca_overall = fit_pca_function(data_wide_PCA) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "Erie"))
pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "Chesapeake"))
```

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: Erie",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: Chesapeake",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

@@

data_scaled2 = 
  data_long %>% 
  group_by(region, site, transect, horizon, name) %>% 
  dplyr::summarise(value = mean(value)) %>% 
  left_join(data_scaling_max_min) %>% 
  mutate(scaled = (value - min) / (max - min)) %>% 
  dplyr::select(region, site, transect, horizon, name, scaled) %>%
  pivot_wider(names_from = "transect", values_from = "scaled") %>% 
  mutate(correction = case_when(upland < wetland ~ 1,
                                upland > wetland ~ -1)) %>% 
  pivot_longer(cols = c(upland, wetland, transition), names_to = "transect", values_to = "scaled") %>% 
  mutate(scaled = scaled * correction,
         scaled = case_when(transect == "upland" ~ 0,
                            transect == "wetland" ~ 1,
                            TRUE ~ scaled),
         scaled2 = case_when(scaled > 4 ~ 4,
                             scaled < -1 ~ -1,
                             TRUE ~ scaled),
         label = case_when(scaled > 4 | scaled < -1 ~ round(scaled,1))) %>% 
  # order analytes
  mutate(name = factor(name, levels = c("TC", "TN", "TS", "SOM",
                                        "pH", "spConductance",
                                        "WEOC", "DIC",
                                        "Fe", "P",
                                        "Ca", "Mg", "Na", "K", "Al", "CEC",
                                        "NH4N", "NO3N", "Cl", "SO4")),
         site = factor(site, levels = c("CRC", "PTR", "OWC",
                                        "GCW", "MSM", "GWI")))

data_scaled2 %>%
  ggplot(aes(x = scaled2, y = name))+
  geom_segment(aes(x = 0, xend = scaled2, yend = name), 
               color = "black", linetype = "dashed", linewidth = 0.2)+
  geom_segment(aes(x = 0, xend = 1, yend = name), color = "black")+
  geom_point(data = data_scaled2 %>% filter(transect != "transition"),
             color = "black",
             size = 1)+
  geom_point(data = data_scaled2 %>% filter(transect == "transition"), 
             size = 4, aes(color = site), stroke = 1, shape = 21, fill = "white")+
  annotate("text", label = "U", x = 0, y = 21)+
  annotate("text", label = "W", x = 1, y = 21)+
  annotate("text", label = "", x = 1, y = 21.5)+
  xlim(-1, 5)+
  scale_y_discrete(limits = rev)+
  scale_shape_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
                     values = c(21, 22, 24, 21, 22, 24))+
  labs(x = "",
       y = "",
       shape = "transition soils at")+
  facet_wrap(~region)+
  theme_kp()+
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "right",
        legend.background = element_blank()
  )

@@

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

@@

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
#  facet_wrap( ~ name)+
  facet_nested_wrap(~ name + region, ncol = 8,
                    nest_line = element_line(colour = "grey"))+ 
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())
```

@@

```{r scaled_gg2_gwi, fig.height = 8, fig.width = 8}
data_scaled_NEW %>% 
  filter(site == "GWI") %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name + region)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-3, 5)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical",
        axis.text = element_text(size = 16, color = "black"),
        axis.title = element_text(size = 16, face = "bold", color = "black"),
  )+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-5, 3)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical",
        axis.text = element_text(size = 16, color = "black"),
        axis.title = element_text(size = 16, face = "bold", color = "black"),
  )+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-3, 5)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "", title = "Lake Erie")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical",
        plot.title = element_text(face = "bold", size = 18))+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-5, 3)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "", title = "Chesapeake Bay")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical",
        plot.title = element_text(face = "bold", size = 18))+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

@@

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
#  facet_wrap( ~ name)+
  facet_nested_wrap(~ name + region, ncol = 6,
                    nest_line = element_line(colour = "grey"))+ 
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())
```

@@

data_scaled_NEW2 %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 4, aes(shape = site, fill = site), color = "black")+
#  facet_wrap( ~ name)+
  facet_nested_wrap(~ name + region, ncol = 4,
                    nest_line = element_line(colour = "grey"))+ 
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
    scale_fill_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_shape_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c(21, 23, 24, 21, 23, 24))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  scale_y_continuous(breaks = c(0, 0.5, 1), labels = c(0, "" , 1))+
  labs(x = "Transect", y = "Scaled concentration
       ")+
  theme_kp()+
  theme(#axis.text.y = element_blank(),
        panel.grid = element_blank()
        )

@@

```{r v2-scaled_gg2_gwi, fig.height = 8, fig.width = 8, eval=FALSE}
data_scaled_NEW %>% 
  filter(site == "GWI") %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name )+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  scale_y_continuous(breaks = c(0, 0.5, 1), labels = c(0, "" , 1))+
  labs(x = "Transect", y = "Scaled concentration
       ")+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

@@

```{r analytesx-2, fig.height = 4, fig.width = 8, eval=FALSE}
data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
  ggplot(aes(x = transect, y = Fe))+
  geom_violin(aes(fill = region), alpha = 0.6)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable Fe, μg/g")+
  facet_wrap(~region)+
#  scale_fill_manual(values = c("darkblue", "red"))+
  scale_fill_manual(values = pnw_palette("Bay", 2))+
  theme(legend.position = "none")

data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
  ggplot(aes(x = transect, y = P))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable P, μg/g")+
  facet_wrap(~region)+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")
```

@@

  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
    filter(!is.na(value))

@@

wrc_processed %>% 
  reorder_site() %>% 
    filter(kpa_fit >= 0 | kpa_eval >= 0) %>% 
    mutate(region = factor(region, levels = c("Erie", "Chesapeake"))) %>% 
    ggplot(aes(y = water_content_vol_percent, color = transect))+
    geom_line(aes(x = kpa_fit))+
    geom_point(aes(x = kpa_eval), shape = 1)+
    scale_x_log10(labels = scales::comma)+
  scale_color_manual(values = rev(soilpalettes::soil_palette("redox2", 3)))+
  labs(x = "Water potential, kPa",
       y = "Water content, % v/v",
       color = "")+
  #    facet_wrap(~region + site)+
  facet_nested_wrap(region ~ site)+
  theme_kp()
```

@@

elevation = read.csv("1-data/synoptic-elevations.csv", na = "")
elevation_processed = 
  elevation %>% 
  mutate(transect = factor(transect, levels = c("water edge", "wetland", "transition", "upland"))) %>% 
  reorder_site()

gg_wle = 
  elevation_processed %>% 
  filter(region == "Erie") %>% 
  ggplot(aes(x = distance_to_water_m, y = elev_m, color = transect))+
  geom_line(color = "black")+
  geom_point(size = 4)+
  scale_color_manual(values = c("grey", "#2E5894", "#FFB200", "#96001B"),
                     breaks = c("water edge", "wetland", "transition", "upland"))+
  facet_nested_wrap(region ~ site)+
  labs(x = "",
       y = "Elevation, m
       ",
       color = "")+
#  theme(legend.position = c(0.1,0.61))+
  NULL

>>>>

```{r targets}
targets::tar_load(c(data_combined, sample_key))
targets::tar_load_everything()
```

```{r}
reorder_transect = function(dat){
  dat %>% 
    mutate(transect = case_match(transect, "wetland" ~ "marsh", .default = transect),
           transect = factor(transect, levels = c("upland", "transition", "wte", "marsh")))
}

# update transect names in sample_key
sample_key = 
  sample_key %>% 
  reorder_transect()

@@

 # data_combined_subset = 
    data_combined %>% 
    reorder_transect() %>% 
    filter(!transect %in% "wte") %>% 
    force()
}
make_data_wide = function(data_combined_subset, sample_key){

@@

  grp = 
    dat %>% 
    dplyr::select(where(is.character) | where(is.factor)) %>% 
    dplyr::mutate(row = row_number()) %>% 
    right_join(num_row_numbers)

@@

```{r pca_input}
pca_overall = fit_pca_function(dat = data_wide_PCA) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "Erie"))
pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "Chesapeake"))
```

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                     values = pal_transect)+
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: Erie",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                     values = pal_transect)+
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: Chesapeake",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

@@

data_scaled2 = 
  data_long %>% 
  group_by(region, site, transect, horizon, name) %>% 
  dplyr::summarise(value = mean(value)) %>% 
  left_join(data_scaling_max_min) %>% 
  mutate(scaled = (value - min) / (max - min)) %>% 
  dplyr::select(region, site, transect, horizon, name, scaled) %>%
  pivot_wider(names_from = "transect", values_from = "scaled") %>% 
  mutate(correction = case_when(upland < marsh ~ 1,
                                upland > marsh ~ -1)) %>% 
  pivot_longer(cols = c(upland, marsh, transition), names_to = "transect", values_to = "scaled") %>% 
  mutate(scaled = scaled * correction,
         scaled = case_when(transect == "upland" ~ 0,
                            transect == "marsh" ~ 1,
                            TRUE ~ scaled),
         scaled2 = case_when(scaled > 4 ~ 4,
                             scaled < -1 ~ -1,
                             TRUE ~ scaled),
         label = case_when(scaled > 4 | scaled < -1 ~ round(scaled,1))) %>% 
  # order analytes
  mutate(name = factor(name, levels = c("TC", "TN", "TS", "SOM",
                                        "pH", "spConductance",
                                        "WEOC", "DIC",
                                        "Fe", "P",
                                        "Ca", "Mg", "Na", "K", "Al", "CEC",
                                        "NH4N", "NO3N", "Cl", "SO4")),
         site = factor(site, levels = c("CRC", "PTR", "OWC",
                                        "GCW", "MSM", "GWI")))

data_scaled2 %>%
  ggplot(aes(x = scaled2, y = name))+
  geom_segment(aes(x = 0, xend = scaled2, yend = name), 
               color = "black", linetype = "dashed", linewidth = 0.2)+
  geom_segment(aes(x = 0, xend = 1, yend = name), color = "black")+
  geom_point(data = data_scaled2 %>% filter(transect != "transition"),
             color = "black",
             size = 1)+
  geom_point(data = data_scaled2 %>% filter(transect == "transition"), 
             size = 4, aes(color = site), stroke = 1, shape = 21, fill = "white")+
  annotate("text", label = "U", x = 0, y = 21)+
  annotate("text", label = "M", x = 1, y = 21)+
  annotate("text", label = "", x = 1, y = 21.5)+
  xlim(-1, 5)+
  scale_y_discrete(limits = rev)+
  scale_shape_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
                     values = c(21, 22, 24, 21, 22, 24))+
  labs(x = "",
       y = "",
       shape = "transition soils at")+
  facet_wrap(~region)+
  theme_kp()+
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "right",
        legend.background = element_blank()
  )

@@

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "M" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

@@

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
#  facet_wrap( ~ name)+
  facet_nested_wrap(~ name + region, ncol = 8,
                    nest_line = element_line(colour = "grey"))+ 
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "M" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())
```

@@

```{r scaled_gg2_gwi, fig.height = 8, fig.width = 8}
data_scaled_NEW %>% 
  filter(site == "GWI") %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name + region)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "M" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-3, 5)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical",
        axis.text = element_text(size = 16, color = "black"),
        axis.title = element_text(size = 16, face = "bold", color = "black"),
  )+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-5, 3)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical",
        axis.text = element_text(size = 16, color = "black"),
        axis.title = element_text(size = 16, face = "bold", color = "black"),
  )+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-3, 5)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "", title = "Lake Erie")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical",
        plot.title = element_text(face = "bold", size = 18))+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "marsh"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-5, 3)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "", title = "Chesapeake Bay")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical",
        plot.title = element_text(face = "bold", size = 18))+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

@@

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
#  facet_wrap( ~ name)+
  facet_nested_wrap(~ name + region, ncol = 6,
                    nest_line = element_line(colour = "grey"))+ 
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "M" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())
```

@@

data_scaled_NEW2 %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 4, aes(shape = site, fill = site), color = "black")+
#  facet_wrap( ~ name)+
  facet_nested_wrap(~ name + region, ncol = 4,
                    nest_line = element_line(colour = "grey"))+ 
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
    scale_fill_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_shape_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c(21, 23, 24, 21, 23, 24))+
  scale_x_discrete(labels = c("U", "T" , "M" ))+
  scale_y_continuous(breaks = c(0, 0.5, 1), labels = c(0, "" , 1))+
  labs(x = "Transect", y = "Scaled concentration
       ")+
  theme_kp()+
  theme(#axis.text.y = element_blank(),
        panel.grid = element_blank()
        )

@@

```{r v2-scaled_gg2_gwi, fig.height = 8, fig.width = 8, eval=FALSE}
data_scaled_NEW %>% 
  filter(site == "GWI") %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name )+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "M" ))+
  scale_y_continuous(breaks = c(0, 0.5, 1), labels = c(0, "" , 1))+
  labs(x = "Transect", y = "Scaled concentration
       ")+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

@@

```{r analytesx-2, fig.height = 4, fig.width = 8, eval=FALSE}
data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "marsh"))) %>% 
  ggplot(aes(x = transect, y = Fe))+
  geom_violin(aes(fill = region), alpha = 0.6)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable Fe, μg/g")+
  facet_wrap(~region)+
#  scale_fill_manual(values = c("darkblue", "red"))+
  scale_fill_manual(values = pnw_palette("Bay", 2))+
  theme(legend.position = "none")

data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "marsh"))) %>% 
  ggplot(aes(x = transect, y = P))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable P, μg/g")+
  facet_wrap(~region)+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")
```

@@

  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "marsh"))) %>% 
    filter(!is.na(value))

@@

wrc_processed %>% 
  reorder_site() %>% 
  reorder_transect() %>% 
    filter(kpa_fit >= 0 | kpa_eval >= 0) %>% 
    mutate(region = factor(region, levels = c("Erie", "Chesapeake"))) %>% 
    ggplot(aes(y = water_content_vol_percent, color = transect))+
    geom_line(aes(x = kpa_fit))+
    geom_point(aes(x = kpa_eval), shape = 1)+
    scale_x_log10(labels = scales::comma)+
  scale_color_manual(values = rev(soilpalettes::soil_palette("redox2", 3)))+
  labs(x = "Water potential, kPa",
       y = "Water content, % v/v",
       color = "")+
  #    facet_wrap(~region + site)+
  facet_nested_wrap(region ~ site)+
  theme_kp()
```

@@

elevation = read.csv("1-data/synoptic-elevations.csv", na = "")
elevation_processed = 
  elevation %>% 
  mutate(transect = case_match(transect, "wetland" ~ "marsh", .default = transect),
         transect = factor(transect, levels = c("water edge", "marsh", "transition", "upland"))) %>% 
  reorder_site()
